Data Specialization for a Staged Lambda Calculus

### 2015-05-16 paper/semantics.tex
added figure with a smallstep semantics
adding in stuff about new semantics

\subsection{Dynamics}
\label{sec:stagedsemantics}

>>>>

\subsection{(Small Step) Dynamics}
We provide a small-step semantics which evaluates \lang\ terms in two passes,
first reducing only redexes in the first stage portions of code,
then reducing those in second stage portions of code.
Because first- and second-stage code are
interleaved in \lang\ terms, realizing this notion of staged
evaluation requires some care.

The first pass of evaluation comprises three judgements: 
steps $\stepwor e e'$, which asserts that a term $e$ performs some internal reduction to produce $e'$,
where both terms are at world $w$;
doneness $e \done$, which asserts that a term $e$ at world $w$ has no more first stage redeces;
and the auxiliary hoisting $\lift e {e'}$, which will be covered later.

One important feature of our semantics is that it reduces {\em all} first stage redexes
in the first pass, even those that occur underneath second stage binders.
This does not cause evaluation to ``get stuck'' however, 
because first stage code cannot depend on the results of second stage code,
as enforced by our type system.

More formally, the type safety of our dynamics can be stated
with the usual progress and preservation lemmas, annotated with a world:
\\{\bf Progress.~} If $\typeswor e A$, then either $\stepwor e e'$ or $e \done$. 
\\{\bf Preservation.~} If $\typeswor e A$ and $\stepwor e e'$, then $\typeswor e A$. 
Here, $\Gamma$ may be any list of variable bindings at world \bbtwo,
for those second stage variables under which we are evaluating.
Also bear in mind that these lemmas will require minor amendments once hoisting is considered.

\paragraph{Done.} The judgement $e \done$ indicates that a term $e$ contains no first stage redexes.
In the $w = \bbtwo$ case, this means that $e$ contains no first stage code at all; it is a residual.
In the $w = \bbonep$ case, this means a $e$ comprises only tuples, injections, and base value; that is, it is a value.

This second pass is in fact identical to standard monostage lambda calculus evaluation, 
so we will not cover it here.

We show that the type system of \lang\ (specifically, the opaque nature of $\fut$
types) is sufficient to guarantee that our semantics can proceed.

\subsection{(Big Step) Dynamics}
\label{sec:stagedsemantics}
